Food-web structure, stochasticity and response diversity drive temporal stability in complex communities

### 2023-10-19 paper/manuscript.Rmd
Reshape introduction

# IDEA, meeting with Sonia

- Simplifying SEM in main text, move details to appendix
- State diagram for stability - richness relationships along two food-web
  metrics and panels varying response diversity.
- Look at the french data for portefolio and compensatory dynamics effects.
- Mid-October

@@

Global change and its simultaneous increasing stressors constitute a major
threat to biodiversity and ecosystem functioning [@pachauri_climate_2015;
@simmons_refocusing_2021]. Among them, climate change is predicted to lead
dramatically more variable environments [@pachauri_climate_2015]. Understanding
how environmental variability will affect ecosystem
stability is then a timely question.

Biodiversity, in terms of species richness, has been repeatedly shown to drive
higher temporal stability of ecological communities, across various taxa and
ecosystems. Species richness can dampen the effects of perturbations on the
temporal variability of community biomass because more diverse communities
contain species with more diversity of responses to perturbations, which
generates asynchrony in population fluctuations, thereby stabilizing community
biomass fluctuations [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022]. In turn, the effects of human pressures on biotic
homogenization reported in various ecosystems [@buisson_toward_2013;
@muthukrishnan_invasive_2020; @saladin_rapid_2020] is another threat, which
might in turn affect dampens the effects of species richness on ecosystem
stability. However, the joint effects of external stressors such as
environmental variability and internal stressors such as the
loss response diversity has been rarely assessed in temporal stability areas.

The question of stressors can even become complex in multitrophic communities,
where species richness corresponds both to horizontal and vertical diversity.
Food-web structure has been shown to have tremendous effect on community
stability. The degree of generalism of predator was shown to modulate
stability-richness relationships [@thebault_trophic_2005]. The strength and
distribution of interaction strength was also shown to be strong driver of
community stability [@mccann_weak_1998]. Omnivory was also shown to drive higher
stability [@mccann_reevaluating_1997]. The effects of food-web structure on
ecosystem stability has been mostly assessed with small modules [but see
@eschenbrenner_diversity_nodate], and we know very little how food-web structure
affects the relationships between species richness and temporal stability in
complex food-webs.

Both stressors and food-web structure are expected to have different effects on
temporal stability. Stability of community biomass can itself be partitioned in
several components, i.e. population stability and population asynchrony
[@thibaut_understanding_2013; @zhao_biodiversity_2022]. Population stability can
be itself partioned in total biomass and population variability effect.
Population asynchrony can further be partioned in statistical averaging effect
and compensatory dynamics effects [@zhao_biodiversity_2022]. Partitioning
stability allows to better understand how food-web structure and external
factors affect stability. We expect that environmental variability will
primarily affect. For exemple, environmental stochasticity is expected to
primarily affect population variance. Response diversity is expected to act
primarily on compensatory dynamics in the food-webs, as low response diversity
will decrease the effects. Higher connectance and interaction strength are
expected to dampens population variability but to lead to higher synchrony at
the same time.

---
#Understanding the mechanisms that determine temporal
#stability of ecosystem functioning is a long standing
#question in Ecology [@may_will_1972; @angelis_stability_1975;
#@mccann_diversitystability_2000; @tilman_biodiversity_2006;
#@donohue_navigating_2016; @kefi_advancing_2019].
---

In this study, we investigated both the effects of stressors ( environnmental
variability and response diversity) and of food-web structure (connectance,
average interaction strength, average trophic level) on species
richness-stability relationships. To do, we used a bioenergetic food-web model
[@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017] added to
environmental stochasticity.

```{r}
concept_cap <- "Food-web structure, response diversity and enviromental gradient
can modulate stability-richness relationships. (a & b) Food-web structure may
modulate stability-richness relationships additively (a) or interactively (b).
(c) Decrease in response diversity and increase in environmental variability
should dampen stability-richness relationships. While the loss of response
diversity should dampen the effect of species richness on asynchrony, the
increase in environmental variability should decrease primarily population
stability. (d) Increase in productivity should increase species richness
(trophic energy hypothesis) but also dampens the relationship between stability
and species richness (paradox of enrichment). (e) Summary of the framework, how
food-web structure and external perturbations modulate the effects of species
richness on community stability. More precisely, how they affect the different
components of stability: total biomass, population variability, statistical
averaging effects and compensatory dynamics."
```

```{r, eval = FALSE, fig.cap=concept_cap}
knitr::include_graphics(here("report", "figures", "conceptual_figures.pdf"))
```

```{r, eval = FALSE}
tribble(
  ~Type, ~`Food-web structure`, ~`Effect on stability-richness`,
  "Food-web structure", "Connectance", "Strengthen (Miele et al., 2016) / No effect (Thébault & Loreau, 2005)",
  "Food-web structure", "Omnivory", "Strengthen (Mccann et al., 1997)",
  "Food-web structure", "Interaction strength", "Destabilising (Mccann et al., 1998) / Stabilising in
  productive (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Response diversity", "", "Dampens",
  "Productivity", "", "Dampens (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Environmental variability", "", "Dampens",) %>%
kbl(caption = "Predictions of stability-richness relationship from the literature.") %>%
kable_styling()
```

>>>>

---
## IDEA, meeting with Sonia
#- Simplifying SEM in main text, move details to appendix
#- State diagram for stability - richness relationships along two food-web
#  metrics and panels varying response diversity.
#- Look at the french data for portefolio and compensatory dynamics effects.
#- Mid-October
---

@@

Global environmental change including climate, land use and invasive species
constitutes a suite of major threats to biodiversity and ecosystem functioning
[@pachauri_climate_2015; @simmons_refocusing_2021]. A central feature of these
threats is that they are predicted to dramatically increase variability in the
environment that species face [@pachauri_climate_2015] with potential impacts on
the stability of communities and associated impacts on biodiversity and
functioning in these communities. Assessing the impacts of this variability on
communities is challenging for several reasons.

First, there are numerous definitions of community stability depending of the
stressors considered. For example, resistance and resilience are typically used
when a pulse perturbation is considered [@kefi_advancing_2019]. While in natural
settings where press and pulse perturbations are simultaneously present,
community variability measured as the coefficient of variation is typically used
[@thibaut_understanding_2013]. Defined at the community scale, community
variability can be partioned into population stability and asynchrony
[@thibaut_understanding_2013]. Synchronous population fluctuations will amplify
the negative effects of environmental fluctuations on temporal community
stability while compensatory dynamics and portefolio effects (i.e. asynchrony)
will stabilise temporal fluctuations of community biomass [@shoemaker_long_2022;
@zhao_biodiversity_2022]. Environmental forcing (such as food availability and
warming) and land-use intensity was reported to synchronise population
fluctuations [@benton_population_2001; @bluthgen_land_2016; @song_reduced_2015;
@liu_aridity_2023] while habitat degradation tends to decrease population
stability [@olivier_urbanization_2020]. Understanding the effects of global
change on community temporal stability therefore require to account for their
simultaneous effects on community and population scales [@kefi_advancing_2019;
@simmons_refocusing_2021].

Second, the relation between community stability and biodiversity in variable
environment is contingent on species richness per se, on the effects of species
richness on population stability and asynchrony, and ultimately
on the diversity of species responses to enviromental fluctuations [response
diversity, @ives_stability_1999; @elmqvist_response_2003]. Increasing
species richness has been reported to decrease population stability in
experimental grasslands [@tilman_biodiversity_2006; @rosher_identifying_2011],
while increasing population stability in natural settings
[@houlahan_negative_2018] [add mechanisms?]. In turn, higher species richness
often results in higher asynchrony, through portfolio effects (i.e. statistical
averaging effects due to independant species fluctuations) and compensatory
dynamics (due to species interactions and response
diversity of species due to the environment) [@loreau_species_2008;
@thibaut_understanding_2013]. Systematic increase in asynchrony with species
richness is the mechanism that drives reported positive relationships between
community stability and species richness [ref]. Anthropogenic pressures, such as
eutrophication, has been reported to weakens the relationship between species
richness and asynchony in grasslands [@hautier_eutrophication_2014] but such
empiricial evidence remains scarced. However, stressors such as biotic
homogenization [@buisson_toward_2013; @muthukrishnan_invasive_2020;
@saladin_rapid_2020] and environmental forcing that decrease species response
diversity should dampens the relationships between asynchrony and species
richness [@ives_stability_2000; @thibaut_understanding_2013] due to a decrease
in portefolio effects and compensatory effects. [Talk about the effects of
evenness in temporal variability and evenness in abundance? e.g.,
@zhao_biodiversity_2022; @thibaut_understanding_2013]. Because stressors can
affect both community stability per se and how community stability scales with
species richness, improving our understanding require to study both
[@ives_stability_2007; @hautier_anthropogenic_2015].

Third, the structure of species trophic interactions, that determines the
distribution of biomass fluxes through communities, can also control stability.
including levels of generalism, the distribution of interaction strengths and
the prevalence of omnivory, can also influence stability. The degree of
generalism of predator was shown to modulate stability-richness relationships
[@thebault_trophic_2005]. The strength and distribution of interaction strength
was also shown to be strong driver of community stability, such as stronger
interaction strengths can be destabilizing [@may_will_1972; @mccann_weak_1998]. Omnivory was also
shown to drive higher population stability [@mccann_reevaluating_1997].
The effects of food-web structure on ecosystem stability has been mostly
assessed with small modules [but see @eschenbrenner_diversity_nodate], and we
know very little how food-web structure affects the relationships between
species richness and temporal stability in complex food-webs.

Here we deliver inference from modelling of complex biodiverse communities on
how food-web structure (connectance, average interaction strength, average
trophic level), response diversity, and environmental stochasticity combine to
shape species richness-stability relationships.
